---
title: "CIG 2022 Parasite Analysis - Update for Revisions"
author: "Shelby Ziegler"
date: "2024-01-19"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=FALSE, fig.width=14, fig.height=10, out.width = "75%")

rm(list = ls());graphics.off()
library(tidyverse) # For data manipulation and summarization
library(plotrix) #To use standard error calculation
library(cowplot) #For grouping plots
library(scales) #For editing axis scales
library(vegan) #For multivariate analyses
library(rstatix) #For running ANCOVA
library(ggpmisc) #adding equation and r2 to figures
library(viridis)
library(ggpubr)
library(visreg)

alpha=0.05
formula<-y~x
theme_set(theme_classic(base_size=16))
```

Read in dermo and msx prevalence and intensity data 

```{r}
dat<-read.csv("CIGintensityandprev2022.csv")


#calculate co-infection rates
  
dat$coin<-dat$Dermo.prev+dat$MSX.prev
dat$coin<-ifelse(dat$coin>=2, 1, 0)

dat$noin<-dat$Dermo.prev+dat$MSX.prev
dat$noin<-ifelse(dat$noin==0, 1, 0)

dat$MSX.prev[is.na(dat$MSX.prev)] <- 0
dat$coin[is.na(dat$coin)] <- 0
dat$noin[is.na(dat$noin)] <- 0

#levels(dat$Cluster)[levels(dat$Cluster)=="McKay River"] <- "Mackay River"
dat$Cluster<-factor(dat$Cluster, levels=c("Tybee", "Romerly Marsh", "Little Ogeechee", "Liberty-Sunbury", "McIntosh", "McKay River", "Glynn-Jointer Creek", "Camden"))
dat$Site<-factor(dat$Site, levels=c("TB1", "TB2", "TB3", "RM1","RM2", "RM3", "LO1", "LO2", "LO3" ,"SB1", "SB2", "SB3", "MI1", "MI2", "MI3", "MR1","MR2", "MR3", "JC1", "JC2", "JC3", "CM1", "CM2", "CM3"))

```

Calculate and plot proportion of prevalence and coinfection. 

```{r}
#Calculate prevalence for all 4 variables.
prevprop<-dat%>%
  group_by(Cluster, Site)%>%
  summarise(dprop=(sum(Dermo.prev)/n()), mprop=(sum(MSX.prev)/n()), coprop=(sum(coin)/n()), noprop=(sum(noin)/n()))

previn<-dat%>%
  group_by(Cluster, Site)%>%
  summarise(dsum=sum(Dermo.prev), msum=sum(MSX.prev))

previn<-previn[,-c(1,2)]
previn1<-dat%>%group_by(Cluster)%>%summarise(dsum=sum(Dermo.prev), msum=sum(MSX.prev))
previn1<-previn1[,-c(1)]

#Look at independence between Dermo and MSX  
chisq<-chisq.test(table(previn$dsum, previn$msum))
summary(table(previn$dsum, previn$msum))

chisq$observed
chisq$expected

chisqC<-chisq.test(table(previn1$dsum, previn1$msum))
summary(table(previn1$dsum, previn1$msum))

chisqC$observed
chisqC$expected


levels(prevprop$Cluster)[levels(prevprop$Cluster)=="McKay River"] <- "MacKay River"

#write.csv(prevprop, "Prevalence Proportions 2022.csv")
prev1<-prevprop%>%
  group_by()%>%
  summarize(meanpm=mean(dprop), sdpm=sd(dprop),meanhn=mean(mprop), sdhn=sd(mprop), meanco=mean(coprop), sdco=sd(coprop), meanno=mean(noprop), sdno=sd(1-noprop))

chisqX<-chisq.test(table(prev1$meanpm, prev1$meanhn))

prev1$meanpm*prev1$meanhn
prev1$meanco

prev1a<-prev1%>%dplyr::select(meanpm, meanhn, meanco)%>%
  gather(variable, value)

prev1b<-prev1%>%dplyr::select(sdpm, sdhn, sdco)%>%
  gather(sevar, sevalue)

prevX<-cbind(prev1a, prev1b)

prevX$variable<- recode_factor(prevX$variable, meanpm="P. marinus", meanhn="H. nelsoni", meanco="Co-infection")

ggplot(prevX, aes(x=variable, y=value, fill=variable))+
  geom_bar(stat="identity", color="black")+
  geom_errorbar(aes(ymax=value+sevalue, ymin=value-sevalue), width =0.2)+
  ylab("Proportion of infected oysters")+
  xlab("")+
  geom_hline(aes(yintercept=0.5060969), linewidth=2)+
  scale_fill_grey()+
  theme(legend.position="none")+ theme(panel.border = element_rect(colour = "black", fill=NA,  linewidth=1)) +
  theme(panel.grid.major.y = element_blank())
  
  

#Plot dermo prevalence
dermoprevp<-ggplot(prevprop, aes(y=dprop, x=Cluster, fill=Cluster))+
# geom_segment(aes(xend = Cluster), yend = 0, colour = "grey50") +
  geom_point(size = 3, shape=21, color="black") +
  scale_fill_viridis_d() +
  theme(panel.border = element_rect(colour = "black", fill=NA,  linewidth=1)) +
  theme(
    panel.grid.major.y = element_blank(),   # No horizontal grid lines
    legend.position = "none",           # Put legend inside plot area
    legend.justification = c(1, 0.5))+
   theme(axis.text.x=element_text(angle=45,hjust=1)) +
   ylim(0,1)+
  geom_hline(yintercept=0.66, linetype="dashed")+
  scale_x_discrete(limits=rev)+
  labs(y="P. marinus prevalence", x="")+
  theme(axis.text.x=element_blank())

#plot MSX prevalence
msxprevp<-ggplot(prevprop, aes(y=mprop, x=Cluster, fill=Cluster))+
# geom_segment(aes(xend = Cluster), yend = 0, colour = "grey50") +
  geom_point(size = 3, shape=21, color="black") +
  scale_fill_viridis_d() +
  theme(panel.border = element_rect(colour = "black", fill=NA,  linewidth=1)) +
  theme(legend.position = "none",           
    panel.grid.major.y = element_blank())+  # No horizontal grid lines
   # legend.position = c(0.98, 0.25),           # Put legend inside plot area
   # legend.justification = c(1, 0.5))+
  labs(y="H. nelsoni prevalence", x="")+
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  scale_x_discrete(limits=rev)+
   geom_hline(yintercept=0.77, linetype="dashed")+
 ylim(0,1)

#Plot at least 1 infection
noprev<-ggplot(prevprop, aes(y=1-noprop, x=Cluster, fill=Cluster))+
 #geom_segment(aes(xend = Cluster), yend = 0, colour = "grey50") +
  geom_point(size = 3, shape=21, color="black") +
  scale_fill_viridis_d() +
  theme(panel.border = element_rect(colour = "black", fill=NA,  linewidth=1)) +
  theme(legend.position = "none",           
    panel.grid.major.y = element_blank())+  # No horizontal grid lines
   # legend.position = c(0.98, 0.25),           # Put legend inside plot area
   # legend.justification = c(1, 0.5))+
  labs(y="Proportion oysters with at least one parasite", x="")+
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  scale_x_discrete(limits=rev)+
   ylim(0,1)+
  geom_hline(yintercept=0.91, linetype="dashed")
  

prevprop$all<-1-prevprop$noprop

#combine plots 
P1<-cowplot::plot_grid(dermoprevp, msxprevp, ncol=1, rel_heights = c(0.5,0.75))
cowplot::plot_grid(noprev, P1, ncol=2)
```

Calculate average intensity for Dermo and MSX by site and plot 

```{r}
datD<-filter(dat, Dermo.g>0)
datD1<-datD%>%
  group_by(Cluster, Site)%>%
  summarise(Dintensity=mean(Dermo.g), se.in=std.error(Dermo.g))

datD2<-datD1%>%
  group_by()%>%
  summarise(Din=mean(Dintensity), se.in=std.error(Dintensity))

datD1$perc<-datD1$Dintensity/datD2$Din

log(datD2$Din)
log(datD2$se.in)
##issues with jitter 
dip<-ggplot(datD1, aes(x=Cluster, y=log(Dintensity), fill=Cluster))+
#  geom_errorbar(aes(ymax=log(Dintensity)+log(se.in),
 #   ymin=log(Dintensity)-log(se.in)), width=0.1,     position=position_dodge(0.5))+
  geom_point(shape=21,size=3, position=position_dodge(0.5))+
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  scale_x_discrete(limits=rev)+
   theme(panel.border = element_rect(colour = "black", fill=NA,  linewidth=1)) +
  theme(legend.position = "none",           
    panel.grid.major.y = element_blank())+ 
  scale_fill_viridis_d()+
  labs(x="", y="Log(P. marinus DNA/g)")+
  ylim(0,15)+
  geom_hline(yintercept=log(15823.68), linetype="dashed")


datM<-filter(dat, MSX.g>0)

datM1<-datM%>%
  group_by(Cluster, Site)%>%
  summarise(Mintensity=mean(MSX.g), se.in=std.error(MSX.g))

datM2<-datM1%>%
  group_by()%>%
  summarise(Min=mean(Mintensity), se.in=std.error(Mintensity))

datM1$perc<-datM1$Mintensity/datM2$Min

log(datM2$Min)
log(datM2$se.in)

mip<-ggplot(datM1, aes(x=Cluster, y=log(Mintensity), fill=Cluster))+
 #  geom_errorbar(aes(ymax=log(Mintensity)+log(se.in),
  #  ymin=log(Mintensity)-log(se.in)), width=0.1, position=position_dodge(width = 0.5))+
   geom_point(shape=21,size=3, position=position_dodge(width = 0.5))+
   theme(panel.border = element_rect(colour = "black", fill=NA,  linewidth=1)) +
  theme(legend.position = "none",           
    panel.grid.major.y = element_blank())+ 
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  scale_x_discrete(limits=rev)+
  scale_fill_viridis_d()+
  labs(x="", y="Log(H. nelsoni DNA/g)")+
  ylim(0,35)+
  geom_hline(yintercept= 28.76366, linetype="dashed")


cowplot::plot_grid(dip, mip)

anova(lm(log(Dintensity)~Cluster, data=datD1))
TukeyHSD(aov(lm(log(Dintensity)~Cluster, data=datD1)))
anova(lm(log(Mintensity)~Cluster, data=datM1))
```

Macroparasite prevalence  

```{r}
macro<-read.csv("Data/DiseaseoysterMacros.csv")

macro$worm<-ifelse(macro$blister.number>=1, 1, 0)

#levels(macro$Cluster)[levels(macro$Cluster)=="McKay River"] <- "Mackay River"

#dat$Cluster<-factor(dat$Cluster, levels=c("Tybee", "Romerly Marsh", "Little Ogeechee", "Liberty-Sunbury", "McIntosh", "Mackay River", "Glynn-Jointer Creek", "Camden"))
datmac<-left_join(dat, macro, by="Oyster.ID")

#blister worm by cluster
worm1<-glm(worm~Cluster,family=binomial(link="logit"), data=datmac)
summary(worm1)
drop1(worm1, .~., test = "Chisq")
#pea crab prev by cluster
pc1<-glm(pea.crab~Cluster,family=binomial(link="logit"), data=datmac)
summary(pc1)
drop1(pc1, .~., test = "Chisq")

#Calculate prevalence for pea crabs and blister worm
prevmac<-datmac%>%
  group_by(Cluster, Site)%>%
  summarise(wprop=(sum(worm)/n()), pcprop=(sum(pea.crab)/n()), sprop=(sum(sponge.perc)/n()))

prevmac1<-datmac%>%
  summarise(wprop=(sum(worm)/n()), pcprop=(sum(pea.crab)/n()), sprop=(sum(sponge.perc)/n()))

#Plot worm prevalence
wprevp<-ggplot(prevmac, aes(y=wprop, x=Cluster, fill=Cluster))+
 #geom_segment(aes(xend = Cluster), yend = 0, colour = "grey50") +
  geom_point(size = 3, shape=21, color="black") +
  scale_fill_viridis_d() +
  theme(panel.border = element_rect(colour = "black", fill=NA,  linewidth=1)) +
  theme(
    panel.grid.major.y = element_blank(),   # No horizontal grid lines
    legend.position = "none",           # Put legend inside plot area
    legend.justification = c(1, 0.5))+
   theme(axis.text.x=element_text(angle=45,hjust=1)) +
   ylim(0,1)+
  #geom_hline(yintercept=0.5, linetype="dashed")+
  scale_x_discrete(limits=rev)+
  labs(y="Blister worm prevalence", x="")

#Plot pea crab prevalence
pcprevp<-ggplot(prevmac, aes(y=pcprop, x=Cluster, fill=Cluster))+
 #geom_segment(aes(xend = Cluster), yend = 0, colour = "grey50") +
  geom_point(size = 3, shape=21, color="black") +
  scale_fill_viridis_d() +
  theme(panel.border = element_rect(colour = "black", fill=NA,  linewidth=1)) +
  theme(
    panel.grid.major.y = element_blank(),   # No horizontal grid lines
    legend.position = "none",           # Put legend inside plot area
    legend.justification = c(1, 0.5))+
   theme(axis.text.x=element_text(angle=45,hjust=1)) +
   ylim(0,1)+
  #geom_hline(yintercept=0.5, linetype="dashed")+
  scale_x_discrete(limits=rev)+
  labs(y="Pea crab prevalence", x="")

#Plot pea crab prevalence
bsprevp<-ggplot(prevmac, aes(y=sprop, x=Cluster, fill=Cluster))+
 #geom_segment(aes(xend = Cluster), yend = 0, colour = "grey50") +
  geom_point(size = 3, shape=21, color="black") +
  scale_fill_viridis_d() +
  theme(panel.border = element_rect(colour = "black", fill=NA,  linewidth=1)) +
  theme(
    panel.grid.major.y = element_blank(),   # No horizontal grid lines
    legend.position = "none",           # Put legend inside plot area
    legend.justification = c(1, 0.5))+
   theme(axis.text.x=element_text(angle=45,hjust=1)) +
   ylim(0,1)+
  #geom_hline(yintercept=0.5, linetype="dashed")+
  scale_x_discrete(limits=rev)+
  labs(y="Boring sponge prevalence", x="")

cowplot::plot_grid(wprevp, pcprevp, bsprevp, ncol=3)


w1<-glm(worm~Site,family=binomial(link="logit"), data=datmac)
summary(w1)
drop1(w1, .~., test = "F")
r.squaredGLMM(w1)
with(summary(w1), 1 - deviance/null.deviance)

pc1<-glm(pea.crab~Site,family=binomial(link="logit"), data=datmac)
summary(pc1)
drop1(pc1, .~., test = "Chisq")
with(summary(pc1), 1 - deviance/null.deviance)
```

Add in environmental data 

```{r}
env<-read.csv('Data/ReefCharacteristics-2023-08-07.csv') 

env<-env%>%
  dplyr::select(Site, Latitude, Rugosity, Salinity..ppt., Temp...C.)


temp<-env%>%
  group_by()%>%
  summarize(meant=mean(Temp...C.), sdt=sd(Temp...C.))


datlat<-left_join(datmac, env, by="Site")


oy5<-read.csv("Data/CIGOysterAllCombinedWideFormat.csv")
oy5$mudcrab.density<-oy5$large.mudcrab.density+oy5$Small.mudcrab.density
oy5<-oy5%>%
  dplyr::select(Site, Total.oyster.density, Marsh.water.ratio, Nearest.Inlet,Nearest.road,Neartest.Armored.Shoreline, Petro.density, mudcrab.density)

datlat1<-left_join(datlat, oy5, by="Site")
###Physical data 
enviro<-read.csv("Data/DNRyear&5year.csv")
enviro<-enviro[,-1]


#select out spring variables and the 5 year temp and salinity means and vars.
datenv1<-enviro%>%dplyr::select(Site, fullT.mean.x,fullS.mean, fullT.var, fullS.var, Spring.DO.mg.L_mean, Spring.DO.mg.L_var, Spring.pH_mean, Spring.pH_var, Spring.Temp..C_mean, Spring.Temp..C_var, Spring.Salinity.ppt_mean, Spring.Salinity.ppt_var)


fulldat<-left_join(datlat1, datenv1, by="Site")
fulldat<-fulldat[,-1]

fulldat$Dermo.g[is.na(fulldat$Dermo.g)] <- 0
fulldat$MSX.g[is.na(fulldat$MSX.g)] <- 0
```

Prep data for mixed models
```{r}
library(lmerTest)
library(MuMIn)
library(lme4)

fulldat$Cluster<-as.factor(fulldat$Cluster)
fulldat$Site<-as.factor(fulldat$Site)

#Remove 0 intensity values.
fulldatD<-fulldat%>%
  filter(Dermo.g>0)

fulldatM<-fulldat%>%
  filter(MSX.g>0)
```

Look at correlation matrices for variables

```{r}
envcor<-fulldat%>%dplyr::select(Salinity..ppt., Temp...C., fullT.mean.x,fullS.mean, fullT.var, fullS.var, Spring.DO.mg.L_mean, Spring.DO.mg.L_var, Spring.pH_mean, Spring.pH_var, Spring.Temp..C_mean, Spring.Temp..C_var, Spring.Salinity.ppt_mean, Spring.Salinity.ppt_var)

envcor1<-cor(envcor)
corrplot::corrplot(envcor1, method ='number',type = "upper",
         tl.col = "black", tl.srt = 45)

#Look at corplot
allcor<-fulldat%>%
  dplyr::select(Salinity..ppt., Temp...C., Length, Rugosity, Total.oyster.density, mudcrab.density, Nearest.Inlet, Nearest.road, Marsh.water.ratio)

corrplot::corrplot(cor(allcor), method ='number',type = "upper",
         tl.col = "black", tl.srt = 45)


```


Run one full model for dermo presence, msx presence, dermo intensity, msx intensity
Full models

```{r}
library(lme4)
library(MuMIn)
#Run bio variables for Dermo prev
bioD<-glmer(Dermo.prev~Length+scale(mudcrab.density)+scale(Total.oyster.density)+scale(Rugosity)+(1|Site), family=binomial(link="logit"), data=fulldat, na.action=na.fail)

dredge(bioD,"sd", extra = c("r.squaredGLMM", F = function(x)
summary(x)$fstatistic[[1]]))

#Run physical variables.
physD<-glmer(Dermo.prev~Length+scale(Temp...C.)+scale(Marsh.water.ratio)+scale(Nearest.Inlet)+scale(Salinity..ppt.)+(1|Site), family=binomial(link="logit"), data=fulldat, na.action=na.fail)

dredge(physD,"sd", extra = c("r.squaredGLMM", F = function(x)
summary(x)$fstatistic[[1]]))

#Full model based on phys and bio variable models 
fullD<-glmer(Dermo.prev~Length+scale(Temp...C.)+scale(mudcrab.density)+scale(Marsh.water.ratio)+(1|Site), family=binomial(link="logit"), data=fulldat, na.action=na.fail)
summary(fullD)
r.squaredGLMM(fullD)

dredge(fullD,"sd", extra = c("r.squaredGLMM", F = function(x)
summary(x)$fstatistic[[1]])) # marsh:water, length, mud crab + temp 


par(mfrow = c(2, 2))
visreg(fullD, scale="response", rug=2, points=list(cex=1.4, lwd=0.1, col="black", pch=21), partial=T, ylab="P(P. marinus presence)")
par(mfrow = c(1, 1))


## Run models for MSX prev
bioM<-glmer(MSX.prev~Length+scale(mudcrab.density)+scale(Total.oyster.density)+scale(Rugosity)+(1|Site), family=binomial(link="logit"), data=fulldat, na.action=na.fail)

dredge(bioM,"sd", extra = c("r.squaredGLMM", F = function(x)
summary(x)$fstatistic[[1]]))

#Run physical variables.
physM<-glmer(MSX.prev~Length+scale(Temp...C.)+scale(Marsh.water.ratio)+scale(Nearest.Inlet)+scale(Salinity..ppt.)+(1|Site), family=binomial(link="logit"), data=fulldat, na.action=na.fail)

dredge(physM,"sd", extra = c("r.squaredGLMM", F = function(x)
summary(x)$fstatistic[[1]]))

#Full model based on phys and bio variable models 
fullM<-glmer(MSX.prev~Length+scale(Temp...C.)+scale(mudcrab.density)+scale(Marsh.water.ratio)+scale(Salinity..ppt.)+(1|Site), family=binomial(link="logit"), data=fulldat, na.action=na.fail)
summary(fullM)
r.squaredGLMM(fullM)

dredge(fullM,"sd", extra = c("r.squaredGLMM", F = function(x)
summary(x)$fstatistic[[1]])) # marsh:water, length, mud crab + temp 


par(mfrow = c(3, 2))
visreg(fullM, scale="response", rug=2, points=list(cex=1.4, lwd=0.1, col="black", pch=21), partial=T, ylab="P(P. marinus presence)")
par(mfrow = c(1, 1))



#run full Dermo.in minus high correlations
## Run models for MSX prev
bioDI<-lmer(log(Dermo.g)~Length+scale(mudcrab.density)+scale(Total.oyster.density)+scale(Rugosity)+(1|Site), REML=F, data=fulldatD, na.action=na.fail)

dredge(bioDI,"sd", extra = c("r.squaredGLMM", F = function(x)
summary(x)$fstatistic[[1]]))

#Run physical variables.
physDI<-lmer(log(Dermo.g)~Length+scale(Temp...C.)+scale(Marsh.water.ratio)+scale(Nearest.Inlet)+scale(Salinity..ppt.)+(1|Site), REML=F, data=fulldatD, na.action=na.fail)

dredge(physDI,"sd", extra = c("r.squaredGLMM", F = function(x)
summary(x)$fstatistic[[1]]))

#Full model based on phys and bio variable models 
fullDI<-lmer(log(Dermo.g)~scale(Temp...C.)+scale(Marsh.water.ratio)+scale(Salinity..ppt.)+(1|Site), REML=F, data=fulldatD, na.action=na.fail)
summary(fullDI)
r.squaredGLMM(fullDI)

dredge(fullDI,"sd", extra = c("r.squaredGLMM", F = function(x)
summary(x)$fstatistic[[1]])) # marsh:water, length, mud crab + temp 


par(mfrow = c(1, 3))

visreg(fullDI, scale="response", points=list(cex=1.4, lwd=0.4, col="black", pch=21), partial=T, ylab="P(P. marinus intensity)")

par(mfrow = c(1, 1))



## Run models for MSX prev
bioMI<-lmer(log(MSX.g)~Length+scale(mudcrab.density)+scale(Total.oyster.density)+scale(Rugosity)+(1|Site), REML=F, data=fulldatM, na.action=na.fail)

dredge(bioMI,"sd", extra = c("r.squaredGLMM", F = function(x)
summary(x)$fstatistic[[1]]))

#Run physical variables.
physMI<-lmer(log(MSX.g)~Length+scale(Temp...C.)+scale(Marsh.water.ratio)+scale(Nearest.Inlet)+scale(Salinity..ppt.)+(1|Site), REML=F, data=fulldatM, na.action=na.fail)

dredge(physMI,"sd", extra = c("r.squaredGLMM", F = function(x)
summary(x)$fstatistic[[1]]))

#Full model based on phys and bio variable models 
fullMI<-lmer(log(MSX.g)~Length+scale(mudcrab.density)+scale(Salinity..ppt.)+(1|Site), REML=F, data=fulldatM, na.action=na.fail)
summary(fullMI)
r.squaredGLMM(fullMI)
dredge(fullMI,"sd", extra = c("r.squaredGLMM", F = function(x)
summary(x)$fstatistic[[1]])) # marsh:water, length, mud crab + temp 


par(mfrow = c(1, 3))

visreg(fullMI, scale="response", points=list(cex=1.4, lwd=0.4, col="black", pch=21), partial=T, ylab="P(H. nelsoni intensity)")

par(mfrow = c(1, 1))



```